### 2015-04-09 src/5_2_Proposal.tex
propuesta ordenada

The relation between the number of cointegration relations and parameters $L$
and $p$ is studied and then a relation between cointegration and performance is
stated.

Figure~\ref{fig:hists} shows the number $r$ of cointegration vectors given by
the Johansen method for different values of the parameters $L$ and $p$, using
four currencies data, i.e., $l=4$, and considering 1000 iterations $it$.
We know that when there is no cointegration vectors, i.e $r=0$, 
there is no cointegration. In the same way, when $r=l$ reveals that no process is I(1) but,
instead, they are all stationary.
The interesting cases of cointegration are those, where $r$ lies strictly
between $0$ and $l$, i.e. $0<r<l$,

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/histCointVectors}
  \caption{Histogram of the number of cointegration vectors using $p=1$. Four
  possible values for $L$ are shown (100, 1900, 6100 and 10700).
  In the cases of $L=100$ and $L=6100$, most of the time there is little
  cointegration (0 or 4 cointegration vectors). However, for $L=1900$ and
  $L=10700$ significant cointegration is found.}
  \label{fig:hists}
\end{figure}

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/MAPEvsCoint-offset21600-p-2-freq-10s}
  \caption{MAPE versus percentage of cointegration. MAPE was obtained from 1000
  iterations considering several possible $L=[100,10000,400]$ and $p=[1,5]$.
  Best $L$ found was 8900.}
  \label{fig:cointvsmape}
\end{figure}

@@

Algorithm ~\ref{alg:AVECM} summarises our proposal. 
The function {\bf get\_best\_params} makes a grid search on the two vector
lists $Ls$ and $ps$ and returns the parameters $L$ and $p$ which maximise
the percentage of cointegration $PC$ (see equation~\ref{eq:pcoint}) for a
pre-defined number of iterations.
Since this search is computational expensive, the number of cointegration
vectors at each iteration is computed in parallel, thus ensuring a response
before new data is available. 
This parallel routine also allows to refine our grid search leading to obtain
improved forecasting performance.
After that optimum $L$ and $p$ parameters are found, VECM is built and used
to forecast next data.

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$ps$: vector with the number of past values range \\
$Ls$: vector with window sizes range ($L<N$) \\
$outit$: Number of out-of-sample iterations($outit < N-L$)
$it$: Number of iterations of testing (in-sample)\\
$of$: Starting point of Testing \\
\ENSURE  $\,$ \\
$\{ \mathbf{y}_{\text{pred}}[1],\dots,\mathbf{y}_{\text{pred}}[N-L]\}$: model predictions 
\FOR { $i =of$ to $of + outit$ }
   \STATE $\mathbf{Y} \gets \mathbf{y}[i-\texttt{max}(Ls):i]$
    \STATE $L,p \gets
    \texttt{get\_best\_params}(Ls,ps,it,Y)$
    \STATE $\mathbf{y}_i \gets \mathbf{y}[i-L:i]$
        \STATE $model = VECM(\mathbf{y}_i, p)$
        \STATE $\mathbf{y}_{\text{pred}}[i] = model.predict()$
\ENDFOR
\end{algorithmic}
\caption{AVECM: Adaptive VECM.}
\label{alg:AVECM}
\end{algorithm}

\section{Background}
\label{sec:background}

%
%
%\subsection{Evaluation methods} \label{sec:evaluation}
%
%Forecast performance was evaluated using different methods. We chose three
%measures that are frequently used:
%\begin{description}
%\item
%{\bf MAPE}, Mean Average Percent Error, which presents forecast errors as a
%percentage:
%\begin{equation}\label{eq:MAPE}
%\text{MAPE} = \frac{1}{N} \sum_{t=1}^{N} 
%\frac{\left|\mathbf{y}_t-\hat{\mathbf{y}}_t\right|}{\left|\mathbf{y}_t\right|}
% \times 100 
%\end{equation}
%\item
%{\bf MAE}, Mean Average Error, which measures the distance between forecasts and the
%true value.
%\begin{equation}\label{eq:MAE}
%\text{MAE} = \frac{1}{N} \sum_{t=1}^{N} 
%\left| 
%\mathbf{y}_t-\hat{\mathbf{y}}_t
%\right| 
%\end{equation}
%\item
%{\bf RMSE}, Root Mean Square Error, also measures the distance between forecasts
%and the true values but, unlike MAE, large deviations from the true value have a
%large impact on RMSE due to squaring forecast error.
%\begin{equation}\label{eq:RMSE}
%\text{RMSE} = \sqrt{
%\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}}
%\end{equation}
%\end{description}
%
%
%
\section{Experimental results}
\label{sec:results}
\subsection{Data} \label{sec:unitroot}
AVECM tests were carried out using four foreign exchange rates all related to
USD: EURUSD, GBPUSD, USDCHF and USDJPY. This data was collected from the free
database Dukascopy which gives access to the Swiss Foreign Exchange marketplace
~\cite{Dukascopy2014}.

>>>>

\subsection{Motivation}
The relation between the number of cointegration relations
and parameters $L$ and $p$ is studied and then a relation between cointegration
and performance is stated.

Figure~\ref{fig:hists} shows the number $r$ of cointegration vectors given by
the Johansen method for different values of the parameters $L$ and $p$, using
four currencies data, i.e., $l=4$, and 1000 considering iterations ($it=1000$).
\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/histCointVectors}
  \caption{Histogram of the number of cointegration vectors using $p=1$. Four
  possible values for $L$ are shown (100, 1900, 6100 and 10700).
  In the cases of $L=100$ and $L=6100$, most of the time there is little
  cointegration (0 or 4 cointegration vectors). However, for $L=1900$ and
  $L=10700$ significant cointegration is found.}
  \label{fig:hists}
\end{figure}
We know that when there is no cointegration vectors, i.e $r=0$, 
there is no cointegration. In the same way, when $r=l$ reveals that no process is I(1) but,
instead, they are all stationary.
The interesting cases of cointegration are those, where $r$ lies strictly
between $0$ and $l$, i.e. $0<r<l$,

@@

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.7\textwidth]{img/MAPEvsCoint-offset21600-p-2-freq-10s}
  \caption{MAPE versus percentage of cointegration. MAPE was obtained from 1000
  iterations considering several possible $L=[100,10000,400]$ and $p=[1,5]$.
  Best $L$ found was 8900.}
  \label{fig:cointvsmape}
\end{figure}

@@

Algorithm ~\ref{alg:AVECM} solves equation \ref{eq:vareq} defining first
parameters $L$ and $p$ using the function {\bf get\_best\_params} which makes a
grid search on the two vector lists $Ls$ and $ps$ and returns the parameters $L$
and $p$ which maximise the percentage of cointegration $PC$ (see
equation~\ref{eq:pcoint}) for a pre-defined number of iterations.  Since this
search is computational expensive, the number of cointegration vectors at each
iteration is computed in parallel, thus ensuring a response before new data is
available.  This parallel routine also allows to refine our grid search leading
to obtain improved forecasting performance.  After that optimum $L$ and $p$
parameters are found, VECM is built and used to forecast next data.

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$ps$: vector with the number of past values range \\
$Ls$: vector with window sizes range ($L<N$) \\
$outit$: Number of out-of-sample iterations($outit < N-L$)
$it$: Number of iterations of testing (in-sample)\\
$of$: Starting point of Testing \\
\ENSURE  $\,$ \\
$\{ \mathbf{y}_{\text{pred}}[1],\dots,\mathbf{y}_{\text{pred}}[N-L]\}$: model predictions 
\FOR { $i =of$ to $of + outit$ }
   \STATE $\mathbf{Y} \gets \mathbf{y}[i-\texttt{max}(Ls):i]$
    \STATE $L,p \gets
    \texttt{get\_best\_params}(Ls,ps,it,Y)$
    \STATE $\mathbf{y}_i \gets \mathbf{y}[i-L:i]$
        \STATE $model = VECM(\mathbf{y}_i, p)$
        \STATE $\mathbf{y}_{\text{pred}}[i] = model.predict()$
\ENDFOR
\end{algorithmic}
\caption{AVECM: Adaptive VECM.}
\label{alg:AVECM}
\end{algorithm}
%
%
%\subsection{Evaluation methods} \label{sec:evaluation}
%
%Forecast performance was evaluated using different methods. We chose three
%measures that are frequently used:
%\begin{description}
%\item
%{\bf MAPE}, Mean Average Percent Error, which presents forecast errors as a
%percentage:
%\begin{equation}\label{eq:MAPE}
%\text{MAPE} = \frac{1}{N} \sum_{t=1}^{N} 
%\frac{\left|\mathbf{y}_t-\hat{\mathbf{y}}_t\right|}{\left|\mathbf{y}_t\right|}
% \times 100 
%\end{equation}
%\item
%{\bf MAE}, Mean Average Error, which measures the distance between forecasts and the
%true value.
%\begin{equation}\label{eq:MAE}
%\text{MAE} = \frac{1}{N} \sum_{t=1}^{N} 
%\left| 
%\mathbf{y}_t-\hat{\mathbf{y}}_t
%\right| 
%\end{equation}
%\item
%{\bf RMSE}, Root Mean Square Error, also measures the distance between forecasts
%and the true values but, unlike MAE, large deviations from the true value have a
%large impact on RMSE due to squaring forecast error.
%\begin{equation}\label{eq:RMSE}
%\text{RMSE} = \sqrt{
%\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}}
%\end{equation}
%\end{description}
%
%
%
\section{Experimental results}
\label{sec:results}
\subsection{Data} \label{sec:unitroot}
AVECM tests were carried out using four foreign exchange rates all related to
USD: EURUSD, GBPUSD, USDCHF and USDJPY. This data was collected from the free
database Dukascopy which gives access to the Swiss Foreign Exchange marketplace
~\cite{Dukascopy2014}.
